---
output: github_document
---

<!-- badges: start -->
<!-- [![R-CMD-check](https://github.com/datasciencecampus/jtstats-r/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/datasciencecampus/jtstats-r/actions/workflows/R-CMD-check.yaml) -->
<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# jtstats

<!-- badges: start -->
<!-- badges: end -->

The goal of jtstats is to enable easy import of the Department for Transport's Journey Times Statistics (JTS).

## Installation

You can install the development version of jtstats from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("datasciencecampus/jtstats-r")
# # Or for local development, uncomment the following::
# remotes::install_local(".")
# devtools::load_all() # or Ctlr+Shift+B
```

## Usage

Load the package as follows:

```{r}
library(jtstats)
```

For the purposes of this README we will also load the tidyverse metapackage:

```{r tidyerse, message=FALSE}
library(tidyverse)
```


To see what tables are available you can browse [the JTS website on gov.uk](https://www.gov.uk/government/statistical-data-sets/journey-time-statistics-data-tables-jts).
Alternatively, you can check the datasets from within R:

```{r example}
dim(jts_tables)
head(jts_tables)
```

As the above output shows, there are `r nrow(jts_tables)` separate tables that constitute the JTS dataset.
JTS tables are divided into different table types:

```{r}
unique(jts_tables$table_type)
```

You can search for specific JTS tables with `lookup_jts_table()`:

```{r}
jts04_gps = lookup_jts_table(type = "jts04", purpose = "GPs")
jts04_gps$table_title
jts04_gps$sheet
```

As an example, our packages allow easy retrieval of data on the average journey time to employment centres (with 100 to 499 jobs) by public transport simply by running the following lines of code:

```{r}
jts_df = get_jts(type = "jts05", purpose = "employment", sheet = 2019)
```


Imagine you're interested in how average travel time to GP services changed between 2017 and 2019.
You can do that as follows:

```{r}
jts_las_gps_2017 = get_jts(type = "jts04", purpose = "GPs", sheet = 2017)
jts_las_gps_2019 = get_jts(type = "jts04", purpose = "GPs", sheet = 2019)
names(jts_las_gps_2017)
```

```{r}
jts_geo = get_jts(type = "jts04", purpose = "GPs", sheet = 2017, geo = TRUE)
names(jts_geo)


jts_geo %>% 
  select(GPPT15pct) %>% 
  plot()
```
